Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
CP

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee.  We
propose a variant of Lazy Tagging and Clearing that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it. 
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow integrity'' vs. ``memory safety'').}}

@@

\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
  This surely needs \emph{some} foreshadowing in the abstract and intro?}
Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

@@

\bcp{I think we are not explicit enough about the sorts of threat situations
  we imagine we are addressing.  A canonical example (that we could spell
  out) is a developer who wants to call out to potentially buggy or
  malicious libraries.  The library code is going to be linked with the
  developer's in the same address space, but the developer wants her private
  data to be protected from both reading/exfiltrating and overwriting...}

@@

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation\bcp{Can we align the C-like snippets with
    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
      at places in both the C and asssembly, making them harder to read
  and eating more space.}}  \ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
so the final return value produced by {\tt main} is wrong.

The black rectangle represents the return address, which is saved by the call
instruction itself, so is not sealed.

@@

\section{Do Micro-Policies Enforce Stack Safety?}
\label{sec:enforcement}

\bcp{``Enforcement'' seems like too broad a title.  What about calling it
  ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
  Safety?''? And what about folding the testing section into this one?  That
would keep related things together, would avoid appearing to make overly
grand claims in either section, and would give us an excuse to write less
about the mechanism here---we only need to write enough about the mechanism
itself that readers can understand our validation case study.}
\leo{Agreed. Took a crack at it.}

\leo{I am not sure we've hit the correct level of detail in this
  section.}\bcp{Agreed!}

In this section we sketch an existing enforcement mechanism, Depth
Isolation from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
formal stack-safety property described in the previous section; in
fact, it enforces a stronger, inductive version of that property which
strengthens the local state encapsulation component by checking for
integrity and confidentiality violations at every step of the callee's
execution trace. We validate that claim by developing a property-based
testing framework that can quickly uncover inconsistencies between the
enforcement mechanism and the stack-safety property.

\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program.  To control this monitor, they use a programming model that
allows fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
In such a system, all values in
memory addresses and registers (including the $\PCname$)
are enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is defined as a set of tags and a
collection of software-defined rules, indexed by machine opcode.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the instruction itself, and produces
one of two outcomes: either the instruction is allowed to execute (and generates
tags for the result of the operation and the new $\PCname$), or the machine
fail-stops with a policy violation.
\citeauthor{pump_oakland2015}\bcp{and PIPE, if there are any references to
  cite} have shown that a wide range of micro-policies can
be defined using this scheme.

Efficient execution of these micro-policies relies on hardware implementations,
such as the PUMP architecture~\citep{pump:asplos2015}.  Tags are represented
as word-size bit vectors, stored separately from the regular memory and registers.
The hardware incorporates a rule cache
to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
the hardware traps to a software handler (running in a privileged context or
on a co-processor) to compute the rule result. To obtain adequate performance,
it is important to design micro-policies so that they hit in the cache as
much as possible. Thus, practical policies maintain a small working set of
distinct tags.
%% %
%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
%% policies}

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

Observable properties allow us to defer enforcement until a property
violation would become visible, as in lazy policies. But by formalizing
the connection to the eager policy, we can identify when lazy policies
miss violations that can become visible. The Lazy Tagging and Clearing
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity.

Lazy Tagging and Clearing can be repaired if, instead of tagging each function
activation with its depth in the stack, we generate a fresh activation
identifier on each call, which prevents a program from exploiting stale
tags from previous activations. A related mechanism was explored in the
Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
associates a unique activation identifier to each function, which is
however shared by all activations of the same function in order to obtain
good cacheability of tag rules, but that policy allows leaks between
instances of the same function.

@@

\paragraph{Protecting the Stack with Micro-Policies}
%
\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
attacks against stack data are in scope (but not side channels or hardware
attacks), and study the protection of stack data through three families of
micro-policies that tag stack objects with a pair of frame and object identifiers
used to validate accesses to the stack: Return Address Protection (which
prevents an adversary from overwriting designated return addresses), Static
Authorities (which only allows the code of a function to access the stack frames
of its own dynamic instances), and Depth Isolation (described in
\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
temporal locality of stack memory and local call graphs, as well as information
generated by the compilation toolchain, to strike various balances between
precision and cacheability, evaluated through benchmarks that demonstrate
limited performance overhead. In addition to the baseline ``eager'' policies,
they propose a number of lazy optimizations, notably Lazy Tagging and Lazy
Clearing, discussed in \cref{sec:lazy}, with associated improvements in
performance.

We have evaluated Depth Isolation, Lazy Tagging, and Lazy Clearing against
our properties, finding that Depth Isolation does indeed enforce stepwise
stack safety and therefore also observable stack safety. We also found that
neither lazy micro-policy actually does enforce stack safety, and proposed
a fix.

@@

Heap safety, like stack safety, may be framed as a security property
in the form of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
be ``Azedevo et al.''}\fi. Just
as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper
we aim to do the same for the stack.\bcp{Let's describe their work in a bit
  more detail.}
%
Their model describing safety of shared stack-allocated
objects\bcp{??} extends naturally to heap safety, complicated by the fact that
these objects are not deallocated by returns\bcp{??}, and therefore a
function's privilege may increase or decrease after its entry
point. For instance, if a callee allocates a heap object and returns
the pointer to its caller, that object's addresses become accessible
in contradiction to a contour computed at the caller's entry. So such
an extension must enforce trace confidentiality and integrity
properties separately on continuous segments of each call, with
contours computed at each crossing between caller and callee.  \bcp{Didn't
  understand last sentence.}

>>>>

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee.  We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}

@@

\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
  This surely needs \emph{some} foreshadowing in the abstract and
  intro?}\bcp{Yes!}
Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

@@

\bcp{I think we are not explicit enough about the sorts of threat situations
  we imagine we are addressing.  A canonical example (that we could spell
  out) is a developer who wants to call out to potentially buggy or
  malicious libraries.  The library code is going to be linked with the
  developer's in the same address space, but the developer wants her private
  data to be protected from both reading/exfiltrating and
  overwriting... More generally, we can talk about the fact that programmers
  use simple reasoning principles to think about their programs (e.g., local
  variables don't change across function calls), and life is better when
  these principles are actually true...}

@@

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation\bcp{Can we align the C-like snippets with
    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
      at places in both the C and asssembly, making them harder to read
  and eating more space.}\bcp{I'm just talking about removing the space
  above "main" in the left column and moving f down a bit.}}
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
so the final return value produced by {\tt main} is wrong.
%
(The black rectangle represents the return address, which is saved by the call
instruction itself, so it is not sealed.)

@@

\section{Case Study: Do Micro-Policies Enforce Stack Safety?}
\label{sec:enforcement}

% \bcp{``Enforcement'' seems like too broad a title.  What about calling it
%   ``Evaluating Micro-Policies'' or ``Do Micro-Policies Enforce Stack
%   Safety?''? And what about folding the testing section into this one?  That
% would keep related things together, would avoid appearing to make overly
% grand claims in either section, and would give us an excuse to write less
% about the mechanism here---we only need to write enough about the mechanism
% itself that readers can understand our validation case study.}
% \leo{Agreed. Took a crack at it.}

% \leo{I am not sure we've hit the correct level of detail in this
%   section.}\bcp{Agreed!}

In this section we sketch an existing enforcement mechanism, {\em Depth
Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
formal stack-safety property described in the previous section. (In
fact, it enforces a slightly stronger, inductive version of that property which
strengthens the local state encapsulation component by checking for
integrity and confidentiality violations at every step of the callee's
execution trace.) We validate this claim with a property-based
testing framework that can quickly uncover inconsistencies between the
enforcement mechanism and the stack-safety property.

\subsection*{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
\citeauthor{pump_oakland2015}\bcp{additional citations: the noninterference
  paper, Nick and Andre's papers, ...} show that a wide range of
micro-policies can be defined using this scheme.

Efficient execution of these micro-policies relies on hardware acceleration,
such as the PUMP architecture~\citep{pump:asplos2015}.
The hardware incorporates a rule cache
to allow quick retrieval of rule outputs for mapped inputs. If the cache misses,
the hardware traps to a software handler (running in a privileged context or
on a co-processor) to compute the rule result. To obtain adequate performance,
it is important to design micro-policies so that they hit in the cache as
much as possible. Thus, practical policies need to maintain a small working
set of distinct tags.
%% %
%% \rb{Maybe go into more detail, examples later\ldots or try to segue into stack
%% policies}

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the {\em Depth Isolation} policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

Observable properties allow us to defer enforcement until a property
violation would become visible, as in lazy policies. But by formalizing
the connection to the eager policy, we can identify when lazy policies
miss violations that can become visible. The {\em Lazy Tagging and Clearing}
micro-policy from \citet{DBLP:conf/sp/RoesslerD18} tags each stack slot with the call depth
at which it was written, and enforces that it must be read from the
same depth. However, this allows violations of observable stack safety!
Specifically, a callee might write into the stack and leave tags behind
for a future callee at the same depth to read. This violates confidentiality and,
if the writes were in the caller's frame, integrity.

{\em Lazy Tagging and Clearing} can be repaired if, instead of tagging each function
activation with its depth in the stack, we generate a fresh activation
identifier on each call, which prevents a program from exploiting stale
tags from previous activations. A related mechanism was explored in the
Static Authorities policy of \citet{DBLP:conf/sp/RoesslerD18}, which
associates a unique activation identifier to each function, which is
however shared by all activations of the same function in order to obtain
good cacheability of tag rules, but that policy allows leaks between
instances of the same function.

@@

\paragraph{Protecting the Stack with Micro-Policies}
%
\citet{DBLP:conf/sp/RoesslerD18} consider a standard attacker model where all
attacks against stack data are in scope (but not side channels or hardware
attacks), and study the protection of stack data through three families of
micro-policies that tag stack objects with a pair of frame and object identifiers
used to validate accesses to the stack: {\em Return Address Protection} (which
prevents an adversary from overwriting designated return addresses), {\em Static
Authorities} (which only allows the code of a function to access the stack frames
of its own dynamic instances), and {\em Depth Isolation} (described in
\cref{sec:enforcement}). All these policies exploit various kinds of spatial and
temporal locality of stack memory and local call graphs, as well as information
generated by the compilation toolchain, to strike various balances between
precision and cacheability, evaluated through benchmarks that demonstrate
limited performance overhead. In addition to the baseline ``eager'' policies,
they propose a number of lazy optimizations, notably {\em Lazy Tagging} and
{\em Lazy
Clearing}, discussed in \cref{sec:lazy}, with associated improvements in
performance.

We have evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
  Clearing} against our properties, finding that {\em Depth Isolation} does
indeed enforce stepwise stack safety and therefore also observable stack
safety. We also found that neither lazy micro-policy actually does enforce
stack safety, and proposed a fix.

@@

Heap safety, like stack safety, may be framed as a security property
in the form of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
(or ask arthur how!).}\fi. Just
as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper,
inspired by their work, we aim to do the same for the stack.
%
Their model describing safety of shared stack-allocated
objects\BCP{huh?? they were not talking about stacks} extends naturally to heap safety, complicated by the fact that
these objects are not deallocated by returns\bcp{??}, and therefore a
function's privilege may increase or decrease after its entry
point. For instance, if a callee allocates a heap object and returns
the pointer to its caller, that object's addresses become accessible
in contradiction to a contour computed at the caller's entry. So such
an extension must enforce trace confidentiality and integrity
properties separately on continuous segments of each call, with
contours computed at each crossing between caller and callee.  \BCP{Didn't
  understand very much of that paragraph.}
